Demographic Groups with Stronger Biden Support in the Election Race

### 2024-04-05 paper/paper.qmd
update change

In addition, by leveraging data from the 2022 CES, this paper delves into how variables such as education, income, and gun ownership serve as predictors for electoral choices, specifically the likelihood of voting for Biden or Trump. This analysis not only reveal the individual impact of these factors but also contributes to a broader understanding of the socio-economic and cultural impact of American electoral behavior.

Based on our model, I found that individuals with higher incomes demonstrate a greater propensity to vote for Biden. Additionally, elevated levels of personal gun ownership significantly correlate with an increased likelihood of supporting Biden. Conversely, individuals possessing higher levels of education exhibit a reduced probability of voting for Biden.

@@

ces2022 |> slice_head(n = 10) |>
  kable(booktabs = TRUE, caption = "Brief summary of variables types") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```
After installing `dataverse`[@dataverse], I use `get_dataframe_by_name` to analyze CES. I select four variables that most relevant to our topic, which are `votereg`, `presvote20post`, `educ`, `faminc_new`, `gunown`.  `votereg` represents the voter registration status and only the voter who registered to vote would be considered. Besides, `presvote20post` represents the president that respondents vote in 2020 president election. `educ` means the education level of the respondents. In addition, `faminc_new` accounts for the family's annual income in 2021 year. `gunown` records the personal gun ownership by asking the respondent to answer whether they or anyone in their household own a gun. However, for the raw data, there are six different answers, including choosing ,Joe Biden, Donald Trump, Jo Jorgensen, Howie Hawkins, other, and did not vote for President. I only consider the respond of Joe Biden and Donald Trump. Therefore, when I clean the data, I select the group of Joe Biden and Donald Trump. In addition, there are 6 different education level, ranging from no high school, high school graduate, some college, 2-year, 4-year, and post-grad. I care about all of the education level, so I take all of those types into consideration. Moreover, there are 16 ranges of family income and the differences between each range is 10,000. To simiplied the data, I classifed them into 4 different types. For the income group which Less than $10,000,  $10,000 - $19,999, $20,000 - $29,999, and $30,000 - $39,999 are defined as "Low" income group. For those who range from $40,000 to $79,999 can seem as  "Middle" income group.  $80,000 to $199,999 can be considered as "High" income group. For the family earn $200,000 to $500,000 or more is very high income group.

@@

@@

@@

```{r}
#| label: fig-vote
#| fig-cap: The distribution of presidential preferences, by education, and income level
#| echo: false
#| warning: false
#| message: false

ggplot(ces2022, aes(x = income_group, fill = voted_for)) +
  geom_bar(position = "fill") +
  labs(x = "Income Group", y = "Proportion", fill = "Voted For") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  theme(legend.position = "bottom")

## Measurements

# Model {#sec-model}
## Model set-up

Define $y_i$ is the political preference of the respondent, where $y_i$ equal to 1 indicates a preference for Biden and $y_i$ equal to 0 represent a preference for Trump. The predictors include $\mbox{education}_i$, which indicates the education level of the respondent and $\mbox{income}_i$ is the income level of the respondent. $\mbox{gunownership}_i$ is the state of whether individual or the people he/she knows own a gun.

@@

## Weaknesses and Next Steps

>>>>

In addition, by leveraging data from the 2022 CES, this paper delves into how variables such as education, income, and gun ownership serve as predictors for electoral choices, specifically the likelihood of voting for Biden or Trump. I used a logistic regression model to identify the likelihood. This analysis not only reveal the individual impact of these factors but also contributes to a broader understanding of the socio-economic and cultural impact of American electoral behavior.

Based on our model, I found that individuals with middle incomes demonstrate a greater propensity to vote for Biden. Additionally, elevated levels of personal gun ownership significantly correlate with an increased likelihood of supporting Biden. Conversely, individuals possessing higher levels of education exhibit a reduced probability of voting for Biden.

@@

ces2022 |> slice_head(n = 10) |>
  kable(booktabs = TRUE, caption = "Brief summary of variables types") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```
After installing `dataverse`[@dataverse], I use `get_dataframe_by_name` to analyze CES. I select four variables that most relevant to our topic, which are `votereg`, `presvote20post`, `educ`, `faminc_new`, `gunown`.  `votereg` represents the voter registration status and only the voter who registered to vote would be considered. Besides, `presvote20post` represents the president that respondents vote in 2020 president election. `educ` means the education level of the respondents. In addition, `faminc_new` accounts for the family's annual income in 2021 year. `gunown` records the personal gun ownership by asking the respondent to answer whether they or anyone in their household own a gun. However, for the raw data, there are six different answers, including choosing ,Joe Biden, Donald Trump, Jo Jorgensen, Howie Hawkins, other, and did not vote for President. I only consider the respond of Joe Biden and Donald Trump. Therefore, when I clean the data, I select the group of Joe Biden and Donald Trump. In addition, there are 6 different education level, ranging from no high school, high school graduate, some college, 2-year, 4-year, and post-grad. I care about all of the education level, so I take all of those types into consideration. Moreover, there are 16 ranges of family income and the differences between each range is 10,000. To simiplied the data, I classifed them into 4 different types. For the income group which less than $10,000,  $10,000 - $19,999, $20,000 - $29,999, and $30,000 - $39,999 are defined as "Low" income group. For those who range from $40,000 to $79,999 can seem as  "Middle" income group.  $80,000 to $199,999 can be considered as "High" income group. For the family earn $200,000 to $500,000 or more is very high income group.

@@

@fig-income-vote appears to be a bar chart displaying the number of respondents categorized by their income level, and which of the two candidates—Biden or Trump—they voted for. There are four income categories presented: Low, Middle, High, and Very High, with an additional category labeled "NA" that likely represents respondents whose income level was not available or not specified.

@@

```
In addition, @fig-gun-ownership-vote shows the number of respondents classified by their personal gun ownership status and which presidential candidate they voted for, either Biden or Trump. The categories of gun ownership are: "No one in the household owns a gun," "Personally own a gun," "Someone in the household owns a gun," and "NA" (which likely stands for data not available).

@@

@fig-education-income-vote combines the predictors of both income and education and displays the voting preferences (for Biden in red and for Trump in blue) of respondents broken down by their level of education (some college, post-grad, no high school (HS), high school graduate, 4-year degree, 2-year degree) and income level categories (high, very high, low, middle, and NA for data not available).

## Measurements 
I will interpret how `educ`, `faminc_new`, and `gunown` be measured.  Respondents were asked a series of questions regarding their voting preferences, demographic information, and other relevant factors such as education, income, and gun ownership. The survey conduct by online questionnaires and each variable is correspond to a question.
To be more specific, in terms of `educ`, the respondents are asked to answer "What is the highest level of education you have completed?" and the respondents can choose their education level. The answers are categorized into "no high school," "high school graduate," "some college," "2-year degree," "4-year degree," and "post-grad."  However, the dataset fails to provide option for the respondents who graduate from secondary school or no education at all.
`faminc_new` refer to the annual income for a family. People will be asked "Thinking back over the last year, what was your family's annual income?" The income ranges from less than $10,000 to more than $500,000. Some respondents might unwilling to disclose their actual income range.
`gunown` indicates the status of gun ownership. In the course of gathering data on firearm ownership, the study must account for potential response biases. Specifically, individuals who may possess firearms unlawfully could exhibit a tendency to provide untruthful responses, driven by a fear of legal consideration.

# Model {#sec-model}
## Model set-up
Define $y_i$ is the political preference of the respondent, where $y_i$ equal to 1 indicates a preference for Biden and $y_i$ equal to 0 represent a preference for Trump. The predictors include $\mbox{education}_i$, which indicates the education level of the respondent and $\mbox{income}_i$ is the income level of the respondent. $\mbox{gunownership}_i$ is the state of whether individual or the people he/she knows own a gun.

@@

## Weaknesses and Next Steps
There are a few potential weaknesses for this paper. Firstly, I only conduct a logistic regression analysis on 500 samples for computational efficiency purpose. It is limited to reflect a representative sample, which might lead to a inaccurate conclusion on the population's attitudes and behaviors.

In addition, the model can identify the relationship or correlation between different variables but it does not necessarily means causality. To build up a causal relationship require more experimental design.

Moreover, select logistics model merely might fail to capture the overall interaction between predictors. In order to get a more detailed interpretation, some complex model might be helpful.

For solving this limitation, the next steps can be firstly, expand the sample size. To design a research in longitudinal approach might also helpful for identify causality. Additionally, introduce other models to further ensure the accuracy and interpretability 
will also be useful.
